Quantile Regressions via Method of Moments with multiple fixed effects

### 2023-10-27 article/article.qmd
s
s

Because the indicator function $\mathbb{1}()$ is not differentiable, we impose some smoothing assumptions:

First:

$$\begin{aligned}
E[h_{3,i}] -\tau &= E[\mathbb{1}\left( q(\varepsilon|\tau) x_i'\gamma +x_i'\beta - y_i \geq 0 \right)] - \tau \\
                &= E\left[E[\mathbb{1}\left( q(\varepsilon|\tau) x_i'\gamma +x_i'\beta - y_i \geq 0 \right)|x]\right] - \tau
\end{aligned}
$$

Because we are aiming to identify the conditional quantile function of $y$, and under the assumption of the DGP it follows that:

$$\begin{aligned}
q(y|\tau) &= q(\varepsilon|\tau) x_i'\gamma +x_i'\beta \\
E[\mathbb{1}\left( q(y|\tau) - y_i \geq 0 \right)|x] &= \tau 
\end{aligned}
$$

and asymptotically, we expect that

$$E[\mathbb{1}\left( q(y|\tau) - y_i \geq 0 \right)|x]=
E[\mathbb{1}\left( q(y|\tau) \geq  y_i \right)|x] = \tau = F_{y |X} \big( q(y|\tau) \big) 
$$

but also that

$$F_{y|X} \big( q(y|\tau)) = \tau = F_{\varepsilon}( q(\varepsilon|\tau) )
$$

where $F_{y|X}$ is the conditional cumulative distribution function (CDF) of $y$, and $F_{\varepsilon}$ is the unconditonal CDF of the iid error $\varepsilon$.

Second, based on non-parametric estimation of the CDF, we can approximate $F_{y|X} \big( q(y|\tau))$ as:

$$F_{y|X} \big( q(y|\tau) ) \approx \frac{\sum_{i=1}^N F_{y|X) ( q(y|\tau) - y_i)

>>>>

Because the indicator function $\mathbb{1}()$ is not differentiable, we borrow from the non-parametric literature, and approximate the indicator function with the integral of a kernel density function $I()$. This function $I()$, is monotonic and symetrical function around zero, with a domain over the real numbers, and a range between 0 and 1.

With an arbirarily small bandwidth $h$, this function will approximate the indicator function:

$$
\lim_{h\rightarrow 0} I\left(\frac{z}{h}\right) \approx \mathbb{1}(z\geq 0)
$$

Thus the function $h_{3,i}$ can be approximated as:

$$h_{3,i} \approx I\left( q(\varepsilon|\tau) x_i'\gamma +x_i'\beta - y_i  \right) - \tau$$

Now, we can obtain the Jacobian matrix $\bar G_{3,1}$ as:

$$\begin{aligned}
\bar G_{3,1} &= -\frac{1}{N} \sum \frac{\partial h_{3,i}}{\partial \beta'} \\
             &= -N^{-1} \sum K_h(q(\varepsilon|\tau) x_i'\gamma +x_i'\beta - y_i ) x_i'
\end{aligned}
$$

Under the assumption that we have enough observations within each combination of $x$, and of multiplicative heteroskedasticity, we have:

$$\begin{aligned}
E(K_h(q(\varepsilon|\tau) x_i'\gamma +x_i'\beta - y_i )|X) &= f_{y|X}(q(\varepsilon|\tau) x_i'\gamma +x_i'\beta) \\
&=\frac{1}{x_i'\gamma} f_{\varepsilon}(q(\varepsilon|\tau)) \\
\end{aligned}
$$

where $f_{y|X}$ is the conditional probability density function of $y$ given $X$, and $f_\varepsilon$ is the unconditional distribution of the standardized error. With this, we can rewrite the Jacobian matrix as:

$$\bar G_{3,1} &= -N^{-1} f_{\varepsilon}\big(q(\varepsilon|\tau)\big) \sum  \frac{x_i'}{x_i'\gamma} 
$$

Asymptotically, however, the expression $\sum\frac{a_i}{b_i}$ can be approximated using taylor expansions by $N\frac{\bar a}{\bar b}$.[^6] Thus, we can rewrite the last term as:

[^6:] This approximation will be useful when we consider the estimation of the influence functions.
$$\bar G_{3,1} = - f_{\varepsilon}\big(q(\varepsilon|\tau)\big) \frac{\bar x_i'}{\bar x_i'\gamma} 
$$

The Jacobian for the second matrix $\bar G_{3,2}$ can be derived similarly:

$$\begin{aligned}
\bar G_{3,2} &= -\frac{1}{N} \sum \frac{\partial h_{3,i}}{\partial \gamma'} \\
             &= -N^{-1} \sum K_h(q(\varepsilon|\tau) x_i'\gamma +x_i'\beta - y_i ) q(\varepsilon,\tau) x_i' \\
             &=-N^{-1} \sum f_{y|x}(q(\varepsilon|\tau) x_i'\gamma +x_i'\beta ) q(\varepsilon,\tau) x_i' \\
             &=-N^{-1} f_{\varepsilon}(q(\varepsilon|\tau)) q(\varepsilon,\tau) \sum \frac{x_i'}{x_i'\gamma} \\
             &=- f_{\varepsilon}(q(\varepsilon|\tau)) q(\varepsilon,\tau) \frac{\bar x_i'}{\bar x_i'\gamma}
\end{aligned}
$$

and the Jacobian for the third matrix $\bar G_{3,3}$ is:

$$\begin{aligned}
\bar G_{3,3} &= -\frac{1}{N} \sum \frac{\partial h_{3,i}}{\partial q(\varepsilon|\tau)} \\
 &= -\frac{1}{N} \sum K_h(q(\varepsilon|\tau) x_i'\gamma +x_i'\beta - y_i ) x_i'\gamma \\
 &= -\frac{1}{N} \sum f_{y|X}(q(\varepsilon|\tau) x_i'\gamma +x_i'\beta) x_i'\gamma \\
 &= -\frac{1}{N} \sum f_{\varepsilon}(q(\varepsilon|\tau)) \frac{x_i'\gamma}{ x_i'\gamma} \\
 &= - f_{\varepsilon}(q(\varepsilon|\tau)) 
\end{aligned}
$$

### Influence functions

**Location coefficients**:

$$\lambda_i(\beta) = \bar G^{-1}_{1,1} h_{1,i} = N (X'X)^{-1}(x_i (y_i-x_i'\beta)) = N (X'X)^{-1}(x_i \nu_i)
$$

**Scale Coefficients coefficients**:

$$\begin{aligned}
\lambda_i(\gamma)&=\bar G_{22}^{-1}\Big(h_{2i}-\bar G_{21} \lambda_i(\beta)\big) \\
&=N (X'X)^{-1}(x_i(|\nu_i|-x_i' \gamma))
-  N E(sign(\nu_i))  (X'X)^{-1} x_i R_i \\
&= N(X'X)^{-1} x_i \Big(2 R_i \big[1(R_i \geq 0) - E(1(R_i \geq 0)) \big]  -x_i'\gamma \Big) \\
&= N(X'X)^{-1} x_i (\tilde R_i  -x_i'\gamma )
\end{aligned}
$$
